Numerical analysis of the influence of magnetic field waveforms on the performance of active magnetic regenerators

### 2019-09-17 PaperMagneticProfiles_JBSMSE.tex
Create list of symbols.
Edit AMR model section.
The main edit was to combine explanations on the
magnetic and fluid profiles into a more cohesive section.

@@

@@

In this work, we perform numerical simulations of a previously developed AMR model, varying the profile and geometric parameters. We also couple a model for calculating  the power comsumption of a proposed valve system, a topic understudied in the literature. With this, we can evaluate the influence of the profile on cooling capacity and coefficient of performance.

@@

Simulations were performed using a one-dimensional AMR mathematical model \cite{bib:asme-mce,bib:trevizoli16_perfor_model}. Here we will show a summarized version focusing on the governing equations. This model solves momentum and energy balance equations for the solid and fluid phases, represented by indices 's' and 'f', respectively. The model geometry is depicted in \autoref{fig:amrmodel}, and assumes regenerators composed of monodisperse packed spheres with porosity $\varepsilon$.

@@

\nomenclature[ak]{$K$}{permeability of the porous medium [\si{\meter\square}]}
\nomenclature[ac]{$c\ped{E}$}{Ergun constant of the porous medium}
\nomenclature[ar]{$\rho$}{density [\si{\kg\per\cubic\meter}]}
\nomenclature[be]{$\varepsilon$}{porosity}
\nomenclature[av]{$V$}{velocity [\si{\meter\per\second}]}
\nomenclature[ap]{$P$}{pressure [\si{\pascal}]}
\nomenclature[bu]{$\mu$}{dynamic viscosity [\si{\pascal\second}]}

\noindent and is solved for the time-dependent uniform fluid velocity through the bed.

@@

\begin{equation}
\label{eq:78}
\begin{split}
  \rho\ped{f} c_{p,\mathrm{f}} \left(\varepsilon \diffp{T\ped{f}}{t} + V_z\diffp{T\ped{f}}{z}\right) = & -\hheat\ped{sf} \beta \left(T\ped{f} - T\ped{s}\right) \\
& + \left|V_z\diffp{p}{z}\bigg|_f \right| \\
& + \varepsilon\left(k\ped{f}\ap{eff}  +\rho\ped{f} c_{p,\mathrm{f}} D\ped{ld}\right)\diffp[2]{T\ped{f}}{z} \\
& + \rate{q}\ped{csg}  
\end{split}
\end{equation}

@@

\nomenclature[aq]{$\rate{q}\ped{csg}$}{volumetric casing losses in the AMR model [\si{\watt\per\cubic\meter}]}
\nomenclature[ic]{csg}{regenerator casing}
\nomenclature[ah]{$\hheat$}{heat transfer coefficient [\si{\watt\per\meter\squared\per\kelvin}]}
\nomenclature[aa]{$A\ped{sf}$}{contact surface area between solid and fluid phases in the regenerator [\si{\meter^2}]}
\nomenclature[is]{sf}{relative to the heat transfer between solid and fluid phases in the regenerator}
\nomenclature[azb]{$\beta$}{surface area density [\si{\squared\meter\per\cubic\meter}]}
\nomenclature[ak]{$k\ap{eff}$}{effective thermal conductivity [\si{\watt\per\meter\per\kelvin}]}
\nomenclature[ad]{$D\ped{ld}$}{dispersion term in the AMR model [\si{\meter\square\per\second}]}
\nomenclature[af]{$f$}{cycle frequency [\si{\hertz}]}
\nomenclature[azf]{$\Phi$}{utilization}
\nomenclature[am]{$m$}{mass [\si{\kg}]}

\nomenclature[azt]{$\tau$}{time period [\si{\second}]}
\nomenclature[ac]{$c$}{specific heat [\si{\joule\per\kg\per\kelvin}]}

@@

Initial and boundary conditions, solution methods, convergence analyses, and closure relations for the porous media terms are described in \cite{bib:trevizoli16_perfor_model}. This AMR model solves the above equations for one regenerator operating between given sources temperatures, during one full cycle (hot and cold blows and  magnetization and demagnetization periods), given specified operating conditions (to be discussed later).

@@

The pressure gradient in \autoref{eq:75}is modeled as:

@@

The magnetic profile is modeled by a waveform of magnetic field strength $H$ applied perpendicular to the regenerators, as shown in \autoref{fig:amrmodel}. The magnetic field is assumed uniform throughout the beds. The applied field is corrected from demagnetization effects to yield the effective field inside the regenerators:

\nomenclature[ct]{$\dtad$}{adiabatic temperature variation [\si{\kelvin}]}

\noindent where the adiabatic temperature variation $\dtad$ is calculated from \textcolor{black}{tabulated} experimental \textcolor{black}{data} for magnetocaloric materials as function of temperature and effective field. \cite{bib:trevizoli16_perfor_model}.

n this chapter, all simulations use gadolinium or its allows as the magnetocaloric material. Gadolinium is a benchmark material with a Curie temperature of \SI{290}{\kelvin}. \textcolor{black}{For simplicity, the solid} density is assumed constant at $\rho\ped{s} = \SI{7900}{\kg\per\cubic\meter}$ and the solid \textcolor{black}{thermal} conductivity is set to $k\ped{s} = \SI{10.5}{\watt\per\meter\per\kelvin}$. The specific heat \textcolor{black}{capacity} of \ce{Gd} is calculated as a function of temperature and magnetic field based on experimental \textcolor{black}{data}, using a bi-linear interpolation scheme; more details on the experimental dataset \textcolor{black}{are available} in \cite{bib:trevizoli16_perfor_model}.

In this work, both single- and multilayer regenerators are considered. For the multilayer simulations, alloys of gadolinium and yttrium are used in the form \ce{Gd_{1-x}Y_x}, where $x$ is the yttrium fraction.  \textcolor{black}{This fraction} decreases the Curie temperature  \textcolor{black}{of the} alloy relative to that of pure gadolinium. Due to the lack of experimental  \textcolor{black}{data on the magnetocaloric properties of} \ce{Gd_{1-x}Y_x} alloys at the time \textcolor{black}{this analysis was made}, a simpler approach was used \textcolor{black}{in which} the properties of alloys with {a} low yttrium fraction are \textcolor{black}{identical to those} of pure gadolinium, except for the Curie temperature \textcolor{black}{(which is shifted to lower values)}.

The AMR model considers only one bed and assumes $N\ped{r}$ identical beds experience the same cycle, multiplyind the performance metrics below by this factor

\nomenclature{an}{$N\ped{r}$}{number of regenerators}

\noindent where $\tau$ is the cycle period, and $\tau\ped{HB}$ is the duration of the hot blow, during which the cold end temperature is compared to the cold source temperature.

@@

\nomenclature[azt]{$\tau\ped{B}$}{blow duration [\si{\second}]}

@@

\subsection{Definition of magnetic profiles}
\label{sec:defin-magn-prof}

The magnetic profiles considered in this work were shown in \autoref{fig:profiles}; here their mathematical definitions are presented. The profiles are presented in terms of the flux density $B = \mu_0 H$, where $\mu_0$ is the permeability of free space; the magnetic field $H$ is used in the evaluation of the magnetocaloric effect (cf. \autoref{sec:how-magnetic-profile}).

The instantaneous profile (represented by the subscript ``IT'') and the rectified cosine profile (represented by ``RC'') are defined solely in terms of the \textcolor{black}{(extreme)} values $B\ped{min}$ and $B\ped{min}$:

As previously discussed, a suitable approximation of the instantaneous profile is the \textcolor{black}{\emph{magnetic ramp profile}, shown} in \autoref{fig:ramp}. The so-called \emph{hot cycle}, \textcolor{black}{during which} the MCM is magnetized, occupies the range \(0 \le t < \nicefrac{\tau}{2}\), while the \emph{cold cycle} lies in the range \(\nicefrac{\tau}{2} \le t < \tau\). The magnetic profile oscillates  between a low value \({B}\ped{min}\) and a high value \({B}\ped{max}\), and \textcolor{black}{remains} at each plateau for a period of \(\tau\ped{M}\). The plateaus are balanced and centered at each half-cycle.

>>>>

\makenomenclature

\renewcommand{\nomgroup}[1]{%
\ifthenelse{\equal{#1}{A}}%
    {\item[\textbf{Variables}]}%
{\ifthenelse{\equal{#1}{B}}%
{\item[\textbf{Greek symbols}]}%
    {\ifthenelse{\equal{#1}{I}}%
        {\item[\textbf{Subscripts and Superscripts}]}%
        {\ifthenelse{\equal{#1}{V}}%
            {\item[\textbf{Abbreviations}]}%
                {}%
        }
    }
}}

% \nomenclature[a ]{}{...} produces a regular symbol
% \nomenclature[b ]{}{...} produces a greek symbol
% \nomenclature[i ]{}{...} produces a subscript
% \nomenclature[v ]{}{...} produces an abbreviation
% The letter after a,i, v or c can control the ordering (e.g. for placing greek letter at the end
\renewcommand{\nomname}{List of symbols}%

@@

\printnomenclature

@@

As previuosly explained, we perform numerical simulations of a previously developed AMR model, varying the profile-specific and geometric parameters. We also couple a model for calculating  the power comsumption of a proposed fluid management system, a topic understudied in the literature. The output of this integrated model are cooling capacity, power contributions and the coefficient of performance (COP).

\nomenclature[ac]{COP}{coefficient of performance}

@@

Simulations were performed using a one-dimensional AMR mathematical model, implemented using the Finite Volumes Method \cite{bib:trevizoli16_perfor_model}. Here we will show a summarized version focusing on the governing equations. This model solves momentum and energy balance equations for the solid and fluid phases, represented by indices `s' and `f', respectively. The model geometry is depicted in \autoref{fig:amrmodel}, and assumes regenerators composed of monodisperse packed spheres with porosity $\varepsilon$.

@@

\nomenclature[ak]{$K$}{permeability of the porous medium [\si{\meter\square}]}
\nomenclature[ac]{$c\ped{E}$}{Ergun constant of the porous medium}
\nomenclature[br]{$\rho$}{density [\si{\kg\per\cubic\meter}]}
\nomenclature[be]{$\varepsilon$}{porosity}
\nomenclature[av]{$V$}{velocity [\si{\meter\per\second}]}
\nomenclature[ap]{$P$}{pressure [\si{\pascal}]}
\nomenclature[bu]{$\mu$}{dynamic viscosity [\si{\pascal\second}]}

\noindent and is solved for the time-dependent uniform fluid velocity $V_z$ through the bed.

@@

\begin{equation}
\label{eq:78}
\begin{split}
  \rho\ped{f} c_{p,\mathrm{f}} \left(\varepsilon \diffp{T\ped{f}}{t} + V_z\diffp{T\ped{f}}{z}\right) = & -\hheat\ped{sf} \beta \left(T\ped{f} - T\ped{s}\right) \\
& + \left|V_z\diffp{P}{z}\bigg|_f \right| \\
& + \varepsilon\left(k\ped{f}\ap{eff}  +\rho\ped{f} c_{p,\mathrm{f}} D\ped{ld}\right)\diffp[2]{T\ped{f}}{z} \\
& + \rate{q}\ped{csg}  
\end{split}
\end{equation}

@@

\nomenclature[aq]{$\rate{q}\ped{csg}$}{volumetric casing losses in the AMR model [\si{\watt\per\cubic\meter}]}
\nomenclature[ic]{csg}{regenerator casing}
\nomenclature[ah]{$\hheat$}{heat transfer coefficient [\si{\watt\per\meter\squared\per\kelvin}]}
\nomenclature[aa]{$A\ped{sf}$}{contact surface area between solid and fluid phases in the regenerator [\si{\meter^2}]}
\nomenclature[is]{sf}{relative to the heat transfer between solid and fluid phases in the regenerator}
\nomenclature[bb]{$\beta$}{surface area density [\si{\squared\meter\per\cubic\meter}]}
\nomenclature[ak]{$k\ap{eff}$}{effective thermal conductivity [\si{\watt\per\meter\per\kelvin}]}
\nomenclature[ad]{$D\ped{ld}$}{dispersion term in the AMR model [\si{\meter\square\per\second}]}
\nomenclature[af]{$f$}{cycle frequency [\si{\hertz}]}
\nomenclature[bf]{$\Phi$}{utilization}
\nomenclature[am]{$m$}{mass [\si{\kg}]}

\nomenclature[bt]{$\tau$}{time period [\si{\second}]}
\nomenclature[ac]{$c$}{specific heat [\si{\joule\per\kg\per\kelvin}]}

@@

Initial and boundary conditions, solution methods, convergence analyses, and closure relations for the porous media terms are described in \cite{bib:trevizoli16_perfor_model}. This AMR model solves the above equations for one regenerator operating between given sources temperatures (assuming ideal heat exchangers in contact with the thermal reservoirs), during one full cycle (hot and cold blows and  magnetization and demagnetization periods), given specified operating conditions (to be discussed later).

@@

The pressure gradient in \autoref{eq:75} is modeled as:

@@

@@

\nomenclature[bt]{$\tau\ped{B}$}{blow duration [\si{\second}]}

The magnetic profile is modeled by a waveform of magnetic field strength $H(t)$ applied perpendicular to the regenerators, as shown in \autoref{fig:amrmodel}. The magnetic field is assumed uniform throughout the beds. The applied field is corrected from demagnetization effects to yield the effective field inside the regenerators:

\nomenclature[bt]{$\dtad$}{adiabatic temperature variation [\si{\kelvin}]}

\noindent where the adiabatic temperature variation $\dtad$, a standard measure of the MCE, is calculated from \textcolor{black}{tabulated} experimental \textcolor{black}{data} for magnetocaloric materials as function of temperature and effective field. \cite{bib:trevizoli16_perfor_model}.

The magnetic profiles considered in this work are presented in terms of the flux density $B = \mu_0 H$, where $\mu_0$ is the permeability of free space; the magnetic field $H$ is used in the evaluation of the magnetocaloric effect (cf. Sec.~\ref{sec:how-magnetic-profile}).

The instantaneous profile (represented by the subscript ``IT'') and the rectified cosine profile (represented by ``RC'') are defined solely in terms of the \textcolor{black}{extreme} values $B\ped{min}$ and $B\ped{min}$, and are shown in \autoref{fig:itrc}.

\begin{figure}[!ht]
  \centering
  \includegraphics[width=8cm]{profiles_it_and_rc}
  \caption{Instantaneous ("IT") and rectified cosine ("RC") magnetic profiles}
  \label{fig:itrc}
\end{figure}

A suitable approximation of the instantaneous profile is the \textcolor{black}{\emph{magnetic ramp profile}, shown} in \autoref{fig:ramp}, with finite transition times between the levels of constant magnetization. The magnetic profile oscillates  between a low value \({B}\ped{min}\) and a high value \({B}\ped{max}\), and \textcolor{black}{remains} at each plateau for a period of \(\tau\ped{M}\). The plateaus are balanced and centered at each half-cycle.

In this work, both single- and multilayer regenerators are considered. Single-layer regenerators are composed of gadolinium,  a benchmark material with a Curie temperature of \SI{290}{\kelvin}. \textcolor{black}{For simplicity, the solid} density is assumed constant at $\rho\ped{s} = \SI{7900}{\kg\per\cubic\meter}$ and the solid \textcolor{black}{thermal} conductivity is set to $k\ped{s} = \SI{10.5}{\watt\per\meter\per\kelvin}$. The specific heat \textcolor{black}{capacity} of \ce{Gd} is calculated as a function of temperature and magnetic field based on experimental \textcolor{black}{data}, using a bi-linear interpolation scheme; more details on the experimental dataset \textcolor{black}{are available} in \cite{bib:trevizoli16_perfor_model}.

For the multilayer simulations, alloys of gadolinium and yttrium are used in the form \ce{Gd_{1-x}Y_x}, where $x$ is the yttrium fraction.  \textcolor{black}{This fraction} decreases the Curie temperature  \textcolor{black}{of the} alloy relative to that of pure gadolinium. Due to the lack of experimental  \textcolor{black}{data on the magnetocaloric properties of} \ce{Gd_{1-x}Y_x} alloys at the time \textcolor{black}{this analysis was made}, a simpler approach was used \textcolor{black}{in which} the properties of alloys with {a} low yttrium fraction are \textcolor{black}{identical to those} of pure gadolinium, except for the Curie temperature \textcolor{black}{(which is shifted to lower values)}.

The AMR model considers only one bed and assumes $N\ped{r}$ identical beds experience the same cycle, multiplying the performance metrics below by this factor.

\nomenclature[an]{$N\ped{r}$}{number of regenerators}

@@
